---
title: "Quantified Self Data Analyis"
author: "Peter Petto, for Coursera Reproducible Research Peer Assessment #1"
date: "September 13, 2014"
output: html_document
---
This is an R Markdown file, for the first peer assessment assignment in the John Hopkins' Coursera MOOC, Reproducible Research.

### Loading the Data

This first R code reads in the raw data, displays the first six records and also displays the final six records.

```{r}
rawActivity = read.csv("activity.csv", header = TRUE)
head(rawActivity)
tail(rawActivity)
```

Here is the default summary information for all three raw data variables:

```{r, echo=FALSE}
summary(rawActivity)
```

### Total Number of Steps per Day

Here's a histogram showing the distribution of steps-per-day, for each of the 53 days where step data is available. It is worth noting that there are two dates: 10/2 and 11/15 where the steps-per-day were only 126 and 41 respectively. Although I did not exclude these low outliers, perhaps I should have done so.

```{r, echo=FALSE}
completeActivity <- na.omit(rawActivity)
histInput <- aggregate(completeActivity$steps, by=list(completeActivity$date), FUN=sum)
hist(histInput$x, breaks=9, xlab="Total Steps per Day", main="HISTOGRAM")
```

The mean of total steps per day for this distribution is:

```{r}
mean(histInput$x)
```

And the median of total steps per day for this distribution is:

```{r}
median(histInput$x)
```

### Average Daily Activity Pattern

Here's a timeplot

```{r, echo=FALSE}
timeplotInput <- aggregate(completeActivity$steps, by=list(completeActivity$interval), FUN=mean)
plot(timeplotInput$Group.1, timeplotInput$x, xlab = "Time of Day (military format", ylab = "Average Number of Steps (5min intervals)", main = "Steps, by Time of Day", type ="l")
```

The interval with the greatest number of steps occurs at 835 and contains:

```{r}
max(timeplotInput$x)
```

### Backfilling the Missing Values

The number of missing values in the dataset is:

```{r}
dim(rawActivity[!complete.cases(rawActivity),])[1]
```

I can "backfill" the missing data by filling in the missing value for each time-of-day
interval with the rounded average for that interval for all of the days for which we
do have data, as follows:

```{r}
missingActivity <- rawActivity[!complete.cases(rawActivity),]
```

```{r}
timeplotInput <- cbind(timeplotInput, xr = round(timeplotInput$x))
```

```{r}
missingActivity$steps <- timeplotInput$xr
```

```{r}
imputedActivity <- rbind(completeActivity, missingActivity)
```

Here's a histogram showing the distribution of steps-per-day, for each of the 61 days where step data is available *or imputed.* I did **not** impute data for two dates: 10/2 and 11/15 where the steps-per-day were only 126 and 41 respectively. Perhaps I should have done so.

```{r, echo=FALSE}
hist2Input <- aggregate(imputedActivity$steps, by=list(imputedActivity$date), FUN=sum)
hist(hist2Input$x, breaks=9, xlab="Total Steps per Day", main="HISTOGRAM")
```

The mean of total steps per day for this distribution is:

```{r}
mean(hist2Input$x)
```

And the median of total steps per day for this distribution is:

```{r}
median(hist2Input$x)
```

Because of the method by which I imputed the missing data, the mean did not change, and the median changed only slightly.

### Weekdays vs Weekends

last moves
